Sumayah Alzamil

========================================================

```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)
```


```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using in your analysis in this code
# chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk. This
# prevents the code from displaying in the knitted HTML output. You should set
# echo=FALSE for all code chunks in your file, unless it makes sense for your
# report to show the code that generated a particular plot.

# The other parameters for "message" and "warning" should also be set to FALSE
# for other code chunks once you have verified that each plot comes out as you
# want it to. This will clean up the flow of your report.

library(ggplot2)
library(knitr)
library(dplyr)
library(gridExtra)
library(GGally)
library(scales)
library(RColorBrewer)
library(memisc)

```

```{r echo=FALSE, Load_the_Data}

# Set working Directory
setwd("/Users/sumayah/Desktop/Sumayah/Udacity/Data/R")

# Load the Data
data <- read.csv("wineQualityReds.csv")

names(data)

data$total.acidity <- data$fixed.acidity + data$volatile.acidity

## Set up Plot Themes
theme_set(theme_minimal(10))


```

### Dataset Description - Red Wine

This investigaion is about red wine. It will analyze red wine properties. 

**Data Summary**

```{r echo=FALSE, wine_data_summary}
summary(data)
```

**Data Structure** 

```{r echo=FALSE, wine_data_variable_types}
str(data)
```

# Univariate Plots Section

# X is not necessary so we will remove it
```{r, echo= FALSE, message=FALSE, warning=FALSE}
data <- subset(data, select=-X)
```

```{r echo=FALSE, histogram_of_wine_quality}
ggplot(data = data, mapping = aes(x=quality)) + 
  geom_histogram(binwidth=1)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x = quality), data = data) + geom_bar() + 
  scale_x_continuous(breaks = seq(2, 10, 1))
```

```{r, echo=FALSE, message=FALSE, warning=FALSE, include=FALSE}
plot.hist <- function(xvariable, dataframe, bins) 
  {
  ggplot(dataframe, aes_string(x = xvariable)) + geom_histogram(bins = bins)
}
```

## Fixed Acidity

Let's look at the summary of fixed acidity.
```{r, echo=FALSE, messages=FALSE, warnings=FALSE }
summary(data$fixed.acidity)
```


```{r, echo=FALSE, messages=FALSE, warnings=FALSE }
fah = plot.hist('fixed.acidity', data, 100)  
fah1 = plot.hist('fixed.acidity', data, 100) + 
  scale_x_continuous(limits = c(4, 12), breaks = seq(4, 12, 1))
grid.arrange(fah, fah1)
```


## Volatile Acidity
Let's look at the summary of volatile acidity
```{r, echo=FALSE, messages=FALSE, warnings=FALSE }
summary(data$volatile.acidity)
```


```{r echo=FALSE, warning=FALSE, message=FALSE}
plot.hist('volatile.acidity', data, 100) +
  scale_x_continuous(limits = c(0, 0.75), breaks=seq(0, 0.75, 0.1))
```

## Citric Acid
Let's look at the summary of citric acid
```{r, echo=FALSE, messages=FALSE, warnings=FALSE }
summary(data$citric.acid)
```


```{r, echo=FALSE, message=FALSE, warning=FALSE}
plot.hist('citric.acid', data, 100) +
  scale_x_continuous(limits = c(0, 1.0), breaks = seq(0, 1.0, .1))

```

## Residual Sugar
Let's look at the summary of residual sugar
```{r, echo=FALSE, messages=FALSE, warnings=FALSE }
summary(data$residual.sugar)
```


```{r, echo=FALSE, warning=FALSE, message=FALSE}
rs = plot.hist('residual.sugar', data, 75) 
logrs = plot.hist('residual.sugar', data, 75)  + 
  scale_x_continuous(trans = log10_trans(), breaks=c(0, 1.5, 3, 10, 20, 60))
grid.arrange(rs, logrs)
rsiqr<-8.2 - 1.7 #using 1st & 3rd quartile from summary to find IQR; we'll use this later for outliers
```


## Chlorides



```{r, echo= FALSE, message=FALSE, warning=FALSE}
c = plot.hist('chlorides', data, 75)
c1 = plot.hist('chlorides', data, 75) + scale_x_continuous(limits = c(0, 0.2))
summary(data$chlorides)
grid.arrange(c, c1)
```

```{r, echo= FALSE, message=FALSE, warning=FALSE}
cbp = ggplot(data, aes(y = chlorides, x = 1)) + geom_boxplot(color = 'blue') +
  geom_jitter(aes(x = 1, y = chlorides), alpha = 0.1) + 
  geom_boxplot(color = 'blue', alpha = 1/2) + coord_flip()

grid.arrange(c, cbp)
```


# Univariate Analysis

### What is the structure of your dataset?
Observations of Summary Data:

* The dataset has 1599 observations. 
* Each observation represents a red wine and that red wine has 11 attributes
  1.fixed acidity, 
  2.volatile acidity, 
  3.citric acid, 
  4.residual sugar, 
  5.chlorides, 
  6.free sulfur dioxide, 
  7.total sulfur dioxide, 
  8.density, 
  9.pH, 
  10.sulphates, 
  11.alcohol,
* Red wine has 1 output variable 
  1.red wine quality
  
  

### What is/are the main feature(s) of interest in your dataset?

### What other features in the dataset do you think will help support your \
investigation into your feature(s) of interest?

The main interesting feature  is the wine `quality`. It is the
subjective measurement of how attractive the wine might be to a consumer. The goal here will be to try and correlate non-subjective wine properties with its quality.


### Did you create any new variables from existing variables in the dataset?

### Of the features you investigated, were there any unusual distributions? \
Did you perform any operations on the data to tidy, adjust, or change the form \
of the data? If so, why did you do this?

Many measurements that were clustered close to zero had a positive skew. 
Others such as `pH` and `total.acidity` and `quality` had normal lookingdistributions.

# Bivariate Plots Section
```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_Plots}


## Autocorrelation plots
# ggpairs(data)
        # lower = list(continuous = wrap("points", shape = I('.'))), 
        # upper = list(combo = wrap("box", outlier.shape = I('.'))))
set.seed(666)
ggpairs(data[sample.int(nrow(data),1000),]) + theme(
  axis.ticks = element_blank(),
        axis.text = element_blank())
```


# Bivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. How did the feature(s) of interest c with other features in \
the dataset?

### Did you observe any interesting relationships between the other features \
(not the main feature(s) of interest)?

* There is a relationship between sulphates and wine qualtity, particularly if you disregard extreme sulphate values
* Disregarding measurements where sulphates > 1.0 is the same as disregarding the positive tail of the distribution, keeping just the normal-looking portion
* The relationship is  a week one

### What was the strongest relationship you found?
The strongest relationship between wine properties and wine quality was that of alcohol, chlorides and citric acid. 

# Multivariate Plots Section

```{r,message=FALSE, warning=FALSE, echo=FALSE, Multivariate_Plots}
q1 <- ggplot(aes(x=alcohol, y=chlorides, color=quality), data=subset(data,
          data$chlorides < quantile(data$chlorides, 0.99))) + 
  geom_point(position='jitter') +
  ggtitle("Alcohol Content vs Chlorides and Wine Quality Ratings")

q2 <- ggplot(aes(x=citric.acid, y=pH, color=quality),data=subset(data,
          data$citric.acid < quantile(data$citric.acid, 0.99))) +
  geom_point(position='jitter') +
  geom_smooth() +
  ggtitle("Citric Acid vs pH and Wine Quality Ratings")

grid.arrange(q1,q2)
```

```{r echo=FALSE,message=FALSE, warning=FALSE, Multivariate_Plots2}
q1 <- ggplot(aes(x=density, y=total.acidity, color=quality), data=data) + 
  geom_point(position='jitter') +
  geom_smooth() +
  ggtitle("Density vs Acidity colored by Wine Quality Ratings")

q2 <- ggplot(aes(x=residual.sugar, y=chlorides, color=quality), data=subset(data,
          data$chlorides < quantile(data$chlorides, 0.95))) + 
  geom_point(position='jitter') +
  geom_smooth() +
  ggtitle("Sugar vs Chlorides colored by Wine Quality Ratings")

grid.arrange(q1, q2)
```

# Multivariate Analysis

•	The strongest relationship between wine properties and wine quality was that of alcohol, chlorides and citric acid
•	pH is considered important when assessing wine quality, however the data does not show an appreciable correlation. In fact, any correlations are week ones. 
•	The model built with alcohol, chlorides, citric acid, residual sugar and total acidity featured a low R-squared value of just 0.3. Compared with a model where we just start with a guess of the median wine quality (R-squared of 0.5), the model performs rathe poorly. 
•	Often, the tails of property distributions showed a varied relationship with quality. 



### Talk about some of the relationships you observed in this part of the \
investigation. Were there features that strengthened each other in terms of \
looking at your feature(s) of interest?

### Were there any interesting or surprising interactions between features?

### OPTIONAL: Did you create any models with your dataset? Discuss the \
strengths and limitations of your model.

------

# Final Plots and Summary

### Plot One
```{r echo=FALSE, message=FALSE, warning=FALSE, Plot_One}
ggplot(aes(x=alcohol), data=data) + 
  geom_density(fill=I("#BB2100")) + 
  facet_wrap("quality") + 
  ggtitle("Alcohol Content for Wine Quality Ratings")
```
### Description One
This plot shows the complete picture of the analysis of win quality with two important notes:
1.	Alcohol content has some effect on wine quality rating
2.	The effect is not a strong one


you can almost see in this visualization the distribution shuffle from left to right as wine
ratings increase. 



### Plot Two
```{r echo=FALSE, warning=FALSE, message=FALSE, Plot_Two}
ggplot(aes(x = chlorides, y = alcohol),
       data = subset(data, data$quality != 6)) + 
  geom_point(aes(color = factor(quality))) +
  labs(x = expression(Chlorides~(g/dm^3)),
       y = "Alcohol (percentage of volume)") + 
  scale_color_brewer(type = 'seq',
                     guide = guide_legend(title = 'Wine Quality', 
                                          reverse = T,
                                          override.aes = list(alpha = 1,
                                                              size = 2))) +
  ggtitle("Alcohol by Chlorides for Differing Quality White Wines") + 
  theme(plot.title = element_text(hjust = 0.25, size = 19, face = 'bold'),
        panel.border = element_rect(colour = "black", fill = NA)) 
```

### Description Two
This plot illustrates that there was little to no correlation between wine quality and percentage of volume.  Additionally, in each case, there are outlier ratings that undermine
the general relationship.


### Plot Three
```{r echo=FALSE,message=FALSE, warning=FALSE, Plot_Three}
ggplot(aes(x=citric.acid, y=quality), 
       data=subset(data, data$citric.acid < 
                     quantile(data$citric.acid, 0.99))) + 
  geom_jitter(alpha=2/3) +
  geom_smooth() + ggtitle("SO2")

```


### Description Three
citric acid seems to have the most promeneant effect on wine quality. There is a specific bump at around 0.25 where the wine quality jumps up in the rating. That bump was also present in the univariate distribution plot for citric acid 

When plotted against wine quality in a scatter plot like this, only citric acid
and alcohol displayed a noticeable trend. In each case, the distribution is
slightly skewed towards higher citric acid and higher quality.

------

# Reflection
I don’t have any knowledge about wine before this project, so I didn’t have any expectation or assumption.  Initially, nothing was immediately correlate to be an inherent quality of good wines. However, upon reflection, this is a practical finding. 

I found out after this project that alcohol content and citric acid were the most properties to wine quality.

I was not able to find a strong set of two properties that would consistently be able to predict wine quality with any degree of certainty.

Additionally, having the wine type would be helpful for further analysis. 






